Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-06-27 WSSSPE2_report/WSSSPE2_review.tex
Remove unncessary curly brackets.
undoing some changes

31 submissions were received; all but one used \href{http://arxiv.org}{arxiv}
%arXiv\footnote{\url{http://arxiv.org}}  or figshare\footnote{\url{http://figshare.com}} 
or \href{http://figshare.com}{figshare} to self-publish their papers.

The review process was fairly standard. First, reviewers bid for papers. Then an
automated system matched the bids to determine assignments. After the reviewers
completed their assigned reviews (with an average of 4.9 reviews per paper and
4.1 reviews per reviewer), they used \href{http://easychair.org}{EasyChair} 
%EasyChair\footnote{\url{http://easychair.org/}} 
to record scores on relevance
and comments. Finally, the organizers accessed the information to decide which
papers to associate with the workshop and provided authors with the comments to
help them improve their papers.

@@

The organizers wanted very interactive sessions, with the process of creating
the sessions open to the full program committee, the paper authors, and others
who might attend the workshop. In order to do this, the organizers used 
\href{http://www.well-sorted.org}{Well Sorted},
%Well Sorted\footnote{\url{http://www.well-sorted.org}}
with the following steps:
\begin{enumerate}
%
\item Authors were asked to create Well Sorted ``cards'' for the papers. These
cards have a title (50 characters maximum) and a body (255 characters maximum).
%
\item Authors, members of the WSSSPE program committee, and mailing list subscribers
were asked to sort the cards. Each person dragged the cards, one by one, into
groups. A group could have as many cards as the person wanted it to have, and it
could have any meaning that made sense to that person.
%
\item Well Sorted produced a set of averages of all the sorts, with
various numbers of card clusters.
%
\end{enumerate}

@@

We then moved beyond our training discussion to address funding that encourages
sustainable software -- funding of both people and projects that create a true
system to support sustainability. We called this ``institutionalized
serendipity.'' As science is increasingly reliant on software, one participant
observed that ``software development can have much broader impact than
publishing individual research, but it is not viewed that way.'' Because of this
centrality, one participant mentioned that training ought to be called Science
Carpentry rather than Software Carpentry. We felt we were beginning to see
changes in the research community as a result of the NSF's data management plan
requirements. PIs are thinking more about data and some university libraries are
offering data repositories. We wondered if a software management plan might be
effective. ``How might funding programs need to adapt to reward good software
development practices?'' we asked ourselves. We thought about best practices
such as version control, test harnesses, mailing lists, bug tracking, community
contribution, and reuse where appropriate. We thought about measuring success
through usage statistics (downloads, Altmetrics). ``Should funders demand
reproducibility?'' we wondered. In order for results to be reproduced, software
would need to be carefully curated.

@@

Third that we could change academic culture, rather than worrying about the
systems. This was mostly focused on citations, because they matter for hiring,
promotion, and tenure decisions. The group discussed how we could weigh the
citations within papers better than we now do? How could we identify the five
citations that really matter for a paper, distinguishing them from the related
works and general background that are also cited. Perhaps we could break these
out in the reference list, working with publishers to implement this. Or maybe
we could also break out categories of citations, such as the most important
software used, the previous publication that we are building from, the data that
we actually used, etc. The Moore Foundation's award in data science was given as
an example, asking proposers: What are the five canonical citations that are
most important to your work?~\cite{moore-canonical} This would be a way of
giving credit and assigning importance to these works, differently from how we
just count citations today. A possible action that the group discussed was
conducting a longitudinal study of most useful \{software, data, etc.\} in a
discipline.

@@

\item Paper Policies
\begin{itemize}
	\item Required Sections
	\item Keywords
	\begin{itemize}
		\item Paper provides keywords to help describe software
	\end{itemize}
	\item Papers can be updated when new releases of software are releasesat
	\begin{itemize}
		\item At no extra cost/at significantly reduced cost
	\end{itemize}
\end{itemize}

>>>>

31 submissions were received; all but one used arXiv\footnote{\url{http://arxiv.org}}
or figshare\footnote{\url{http://figshare.com}} to self-publish their papers.

The review process was fairly standard. First, reviewers bid for papers. Then an
automated system matched the bids to determine assignments. After the reviewers
completed their assigned reviews (with an average of 4.9 reviews per paper and
4.1 reviews per reviewer), they used EasyChair\footnote{\url{http://easychair.org/}} 
to record scores on relevance
and comments. Finally, the organizers accessed the information to decide which
papers to associate with the workshop and provided authors with the comments to
help them improve their papers.

@@

The organizers wanted very interactive sessions, with the process of creating
the sessions open to the full program committee, the paper authors, and others
who might attend the workshop. In order to do this, the organizers used 
Well Sorted\footnote{\url{http://www.well-sorted.org}} with the following steps:
\begin{enumerate}
%
\item Authors were asked to create Well Sorted ``cards'' for the papers. These
cards have a title (50 characters maximum) and a body (255 characters maximum).
%
\item Authors, members of the WSSSPE program committee, and mailing list subscribers
were asked to sort the cards. Each person dragged the cards, one by one, into
groups. A group could have as many cards as the person wanted it to have, and it
could have any meaning that made sense to that person.
%
\item Well Sorted produced a set of averages of all the sorts, with
various numbers of card clusters.
%
\end{enumerate}

@@

We then moved beyond our training discussion to address funding that encourages
sustainable software -- funding of both people and projects that create a true
system to support sustainability. We called this ``institutionalized
serendipity.'' As science is increasingly reliant on software, one participant
observed that ``software development can have much broader impact than
publishing individual research, but it is not viewed that way.'' Because of this
centrality, one participant mentioned that training ought to be called Science
Carpentry rather than Software Carpentry. We felt we were beginning to see
changes in the research community as a result of the NSF's data management plan
requirements. PIs are thinking more about data and some university libraries are
offering data repositories. We wondered if a software management plan might be
effective. ``How might funding programs need to adapt to reward good software
development practices?'' we asked ourselves. We thought about best practices
such as version control, test harnesses, mailing lists, bug tracking, community
contribution, and reuse where appropriate. We thought about measuring success
through usage statistics (downloads, altmetrics). ``Should funders demand
reproducibility?'' we wondered. In order for results to be reproduced, software
would need to be carefully curated.

@@

Third that we could change academic culture, rather than worrying about the
systems. This was mostly focused on citations, because they matter for hiring,
promotion, and tenure decisions. The group discussed how we could weigh the
citations within papers better than we now do? How could we identify the five
citations that really matter for a paper, distinguishing them from the related
works and general background that are also cited. Perhaps we could break these
out in the reference list, working with publishers to implement this. Or maybe
we could also break out categories of citations, such as the most important
software used, the previous publication that we are building from, the data that
we actually used, etc. The Moore Foundation's award in data science was given as
an example, asking proposers: What are the five canonical citations that are
most important to your work?~\cite{moore-canonical} This would be a way of
giving credit and assigning importance to these works, differently from how we
just count citations today. A possible action that the group discussed was
conducting a longitudinal study of most useful software, data, etc.\ in a
discipline.

@@

\item Paper Policies
\begin{itemize}
	\item Required Sections
	\item Keywords
	\begin{itemize}
		\item Paper provides keywords to help describe software
	\end{itemize}
	\item Papers can be updated when new releases of software are released
	\begin{itemize}
		\item At no extra cost/at significantly reduced cost
	\end{itemize}
\end{itemize}
